---
title: 'Assignment 7: Stability'
author: "Katheryn Moya and Sarah Lam"
date: "2023-06-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sensitivity)
library(tidyverse)
library(deSolve)
library(ggpubr)
```


## Build Harvest Model 
```{r}

source("~/Documents/MESM Year 2/ESM232/esm232_a7/R/dpredpreyharvest.R")

dpredpreyharvest

```

## Run Model for 200 years 
Start with initial condition of 300 prey and 60 predators, rprey=0.95, alpha=0.01, eff=0.6,pmort=0.4, K=2000.

```{r}
# set initial population size 

initialpop = c(prey = 1, pred = 1)

# come up with first set of parameters

rprey = 0.95
alpha = 0.01
eff = 0.6
pmort = 0.4
K = 200
minprey = 0.6
harv = 0.3

#convert to a dataframe
pars <- cbind.data.frame(rprey, alpha, eff, pmort, K, minprey, harv)

# gets results for 300 years (evaluating every year)
years <-seq(from=1, to=50)

# run model for 50 years using ODE solver

results <- ode(y = Pinitial, times = years, func= dpredpreyharvest, parms = parms)
head(results)

# extract results
res = as.data.frame(results) %>% gather(key="animal", value="pop",-time)

ggplot(as.data.frame(res_test), aes(time, pop , col=animal))+geom_line()
```

4. Run a sensitivity analysis that explores how different hunting levels and different minimum prey populations (before hunting is allowed) are likely to effect the stability of the populations of both predator and prey.

## Sensitivity analysis of different hunting levels and minimum prey populations

To reduce the output dimensionality for our sensitivity analysis, a stability metric must be defined. We decided to define stability as maintaining the prey population above 300 after 50 years

```{r}
# obtain parameters sets (from sobel)
# want to learn about sensitivity to harvest rates and minimum prey population

# set the number of parameters
# all parameters are normally distributed w means as given above and sd of 0.1 of mean
np = 200
sd = 0.1

minprey = rnorm(mean = 300, sd = minprey*sd, n = np)
harv = rnorm(mean = 0.3, sd = harv*sd, n = np)
X1 = cbind.data.frame(minprey = minprey, harv = harv)
            

# repeat to get second set of samples

minprey = rnorm(mean = 300, sd = minprey*sd, n = np)
harv = rnorm(mean = 0.3, sd = harv*sd, n = np)
X2 = cbind.data.frame(minprey = minprey, harv = harv)

# fix negatives since they aren't meaningful
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create sobel object and get sets of parameters for running model 

sens_P = sobolSalt(model = NULL, X1, X2, nboot = 300)

# parameter sets are
head(sens_P$X)

# add names to columns

colnames(sens_P$X) = c("minprey", "harv")

head(sens_P$X)

```

```{r}

### Run ODE with sobol object

# define a wrapper function to do everything we need - run solver and compute metrics - and send back results for each parameter

rprey = 0.95
alpha = 0.01
eff = 0.6
pmort = 0.4
K = 200
minprey = 150
harv = 0.3


pop_wrapper = function (results, pop, pars, times, Pinitial, func){
  pars = list( rprey = rprey, alpha = alpha, eff = eff, pmort = pmort, K = K, minprey = minprey, harv = harv)
  results = ode(y = Pinitial, times = years, func = func, parms = pars)
  colnames(results) = c("time", "prey", "predator")
  
  #get metrics
  
  results = as.data.frame(results)
  stable_prey = mean(results$prey)
  return(list(stable_prey = stable_prey))
}

# use pmap

all_results = as.data.frame(sens_P$X) # order of operations will assume this is the list we are referring to for pmap|


allres = all_results  %>%  pmap(pop_wrapper, results = all_results, Pinitial = Pinitial, times = years, func = dpredpreyharvest)

# extract out results from pmap into a data frame
final_allres = allres %>% map_dfr(`[`,c("stable_prey"))

```


## Establishing a Stability Metric

For the purposes of this assignmnent we decided to define stability as maintaining the prey population above 500 after 50 years

```{r}

```


```{r}
# set parameter list
pars = data.frame(rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000)

# now lets try initial conditions that will be stable
preyi = with(pars, pmort/(eff*alpha)) 
predi = with(pars, rprey/alpha*(1-preyi/K)) 

preyi
predi

# times when you want to evaluate
years = seq(from=1,to=50)
```


```{r}
# establish different harvest rates

lowHrate = 0.015
gps = list(harv = lowHrate, rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000)

# look at the derivative over a range of forest sizes

findstable = data.frame(Pcurr=seq(from=1, to=100, by=5))

findstable$dervHlow= unlist(findstable$Pcurr %>% map(~lotvmodKH(parms=gps, Time=NULL, population=.x) ))


highHrate=0.05
gps = list(harv=highHrate, rprey = 0.95, alpha = 0.01, eff = 0.6, pmort = 0.4, K = 2000)


# plot them all together

tmp = gather(findstable, key="HarvestRate", value="value", -Ccurr)
ggplot(tmp, aes(Ccurr, value, color=HarvestRate))+geom_point()+geom_hline(yintercept = 0, col="black")+
  labs(x="Forest Biomass (kgC)", y="Forest Growth Rate (kgC/year)")

# establish different minimum prey populations (before hunting is allowed)


```


1) build model
2) how stable are things for different minimum pop and hunting
    sensitivity analysis with these two processes
    of what --- 
3) write a function that takes output of model and extracts your definition of stability
 --- if mean pop is above some threshold for X amount of years
 -- or X pop over a certain time
 -- average pop for 5 years
 -- how does this vary with parameters
 --- mean and variance 
4)
5)
6)


